---
title: "FOIA Log Exploration"
author: "Jessica Hirsch"
date: '2022-04-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(tidyverse)
library(here)
library(janitor)
```

```{r}
foia_log_OMB <- read_excel(here::here("Data", "OMB FOIA TRACKING 032719.xlsx" ))
```

```{r}
foia_log_OMB <- clean_names(foia_log_OMB)
```

```{r}
foia_log_OMB <-
  foia_log_OMB %>% 
  rename("date_received" = date_written_request_received_by_agency,
         "date_assigned" = assigned_date,
         "division" = omb_division,
         "date_review" = agency_review_date,
         "is_noncustodial_record" = does_the_request_seek_non_custodial_records_from_dti_omb_or_any,
         "noncustodial_provided_other_agency" = if_non_custodial_records_were_provided_by_another_agency_provid,
         "date_response" = date_of_agency_response_to_requesting_party_under_34_of_policy,
         "date_estimate_sent" = estimate_sent,
         "all_requested_records_sent" = were_all_requested_records_provided_on_date_of_agency_s_respons,
         "date_requested_records_not_provided" = if_all_requested_records_were_not_provided_as_of_date_of_agency,
         "copying_fees" = copying_fees_note_1st_20_pages_free)
```

```{r}
glimpse(foia_log_OMB)
```

How many NA's in each field?

```{r}
foia_log_OMB %>% 
  summarise(across(everything(), ~ sum(is.na(.)))) %>% 
  glimpse()
```

How many Unique Values are there?

```{r}
foia_log_OMB %>% 
  summarise(across(everything(), ~ length(unique(.)))) %>% 
  glimpse()
```

who was assigned foia requests at OMB?

```{r}
foia_log_OMB %>% 
  count(assigned_to)
```

# Strings

## Stringer (`str_`)
```{r}
#str_extract()
#str_remove()
#str_remove_all()
#str_detect()
#str_replace()
#str_replace_all()
```

#Regex
## Who are FOIA logs assigned to?

```{r}
foia_log_OMB <- 
foia_log_OMB %>% 
  mutate(assigned_to = str_remove_all(string = assigned_to,
                                      pattern = "[^A-Za-z]"),
         assigned_to = if_else(is.na(assigned_to), "None", assigned_to))
```

```{r}
foia_log_OMB %>% 
  count(internally_reviewed_by)
```

```{r}
foia_log_OMB <- 
foia_log_OMB %>% 
  mutate(internally_reviewed_by = str_remove_all(internally_reviewed_by,
                                                 "[^A-Za-z]"),
         internally_reviewed_by = if_else(is.na(assigned_to), "None", assigned_to))
```


 #Exploring `Division`
 
 benefits office
 pensions
 PHRST
 HRM
 OSP
 
```{r}
foia_log_OMB <- 
foia_log_OMB %>% 
  mutate(division = case_when(
    str_detect(string = division, pattern = "Dir") ~ "Director's Office",
    str_detect(string = division, pattern = "Ben") ~ "Benefits Office",
    str_detect(string = division, pattern = "Pension") ~ "Pensions",
    str_detect(string = division, pattern = "PHRST") ~ "PHRST",
    str_detect(string = division, pattern = "HR") ~ "HRM",
    str_detect(string = division, pattern = "OSP") ~ "OSP",
    TRUE ~ division
  )) 
```

# `is_noncustodial_record`

```{r}
foia_log_OMB <- 
foia_log_OMB %>% 
  mutate(is_noncustodial_record = case_when(
    str_detect(string = is_noncustodial_record, pattern = "[Nn][Oo]") ~ "No",
    str_detect(string = is_noncustodial_record, pattern = "[Yy][Ee][Ss]") ~ "Yes",
    TRUE ~ is_noncustodial_record
  )) 
```

# `all_requested_records_sent`

Create`all_records_sent_flag`

```{r}
foia_log_OMB <- 
foia_log_OMB %>% 
  mutate(all_requested_records_sent_flag = case_when(
    str_detect(string = all_requested_records_sent, pattern = "^[Nn][Oo]") ~ "No",
    str_detect(string = all_requested_records_sent, pattern = "^[Yy]") ~ "Yes",
    str_detect(string = all_requested_records_sent, pattern = "^[Pp]") ~ "Partial",
    str_detect(string = all_requested_records_sent, pattern = "^[^YyNnPp]") ~ "Other",
    TRUE ~ all_requested_records_sent
  )) 
```

Create `all_requested_records_sent_detail`

```{r}
foia_log_OMB <- 
foia_log_OMB %>% 
  mutate(all_requested_records_sent_detail = str_remove(string = all_requested_records_sent,
                                                 pattern = "^[YyNnPp][\\w,]+[\\s]*"),
       all_requested_records_sent_detail = str_to_sentence(all_requested_records_sent_detail))
```

# `current status`

```{r}
foia_log_OMB <- 
foia_log_OMB %>% 
  mutate(current_status_flag = str_extract(string = current_status,
                                           pattern = "^Closed|^Open|^Inactive"))
```

```{r}
foia_log_OMB <-
  foia_log_OMB %>%
  mutate(current_status_detail = case_when(
           str_detect(string = current_status,
                      pattern = "[Rr]ecords ([Ss]ent|[Ee]mail|[Pp]rovided)|[Ii]nformation [Ss]ent|[Rr]esponse[s]* [Ss]ent|reviewed|reside|responded|fulfilled") ~ "Sent",
           str_detect(string = current_status, pattern = "[Pp]artial") ~ "Partial",
           str_detect(string = current_status, pattern = "[Pp]ossession|no responsive") ~ "Not in possession",
           str_detect(string = current_status, pattern = "[Nn][Oo][Tt]") ~ "Not sent",
           str_detect(string = current_status, pattern = "^Open") ~ "Open",
           str_detect(string = current_status, pattern = "[Ww]ithdrawn|withdrew") ~ "Withdrawn",
           str_detect(string = current_status, pattern = "[Aa]gency|[Ww]eb|internet|DSCYTF|online|DHR|DOL") ~ "Referred elsewhere",
           str_detect(string = current_status, pattern = "[Nn]o res[po]{2}nse") ~ "No response from requestor"
         ))
```


#Factors
## Who was assigned the most requests?

```{r}
foia_log_OMB %>% 
  count(assigned_to) %>% 
  ggplot(aes(x = n,
             y = assigned_to)) +
  geom_col()
```

```{r}
likert <- c(
  "Strongly Agree",
  "Agree",
  "Neutral",
  "Disagree",
  "Strongly Disagree"
)
```

```{r}
sort(likert)
```

```{r}
likert_leveled <- factor(likert, levels = likert)
```

```{r}
sort(likert_leveled)
```

```{r}
set.seed(1231)
sample_survey <-
  data.frame(ID = 1:1000,
             q1 = sample(likert, size = 1000, replace = T),
             q2 = sample(likert, size = 1000, replace = T),
             q3 = round(runif(1000, 1, 5)),
             stringsAsFactors = F)
```

```{r}
sample_survey %>% 
  mutate(q1 = factor(q1, levels = likert),
         q1 = fct_rev(q1)) %>% 
  ggplot(aes(y = q1)) +
  geom_bar()
```

#Forecats (`fct_`)

```{r}
assigned_order <- c("Bert", "Jessica","None")

foia_log_OMB %>% 
  count(assigned_to) %>%
  mutate(assigned_to = factor(assigned_to, levels = assigned_order),
         assigned_to = fct_rev(assigned_to)) %>% 
  ggplot(aes(x = n,
             y = assigned_to)) +
  geom_col()
```


`current_status_detail`

```{r}
foia_log_OMB %>% 
  count(current_status_detail) %>% 
  mutate(current_status_detail = fct_reorder(current_status_detail, n)) %>% 
  ggplot(aes(x = n,
             y = current_status_detail)) +
  geom_col()
```

# `divisions`

```{r}
foia_log_OMB %>% 
  count(division) %>% 
  mutate(division = fct_reorder(division, n)) %>% 
  ggplot(aes(x = n,
             y =division)) +
  geom_col()
```

# Dates

```{r}
foia_log_OMB %>% 
  select(contains("date")) %>% 
  glimpse()
```

```{r}
as.Date("2021-11-01")
as.Date("21/11/1", format = "%y/%m/%d")
as.Date("1/11/21", format = "%d/%m/%y")
as.Date("11/1/2021", format = "%m/%d/%Y")
as.Date("November 1, 2021")
```
```{r}
#library(lubridate)

ymd("2021-11-01")
ymd("21/11/1")
dmy("1/11/21")
mdy("11/1/2021")
mdy("November 1, 2021")
```

